|

### 2019-07-15 inst/manuscript/manuscript.Rmd
Change imputation method to bpca
Change from QRILC from limma package,
to bpca from pcaMethods package.

# set seed for random number generation
set.seed(1)

# reference external R script with protein function lists
knitr::read_chunk('components/protein_funs.R')

@@

# can't impute with so many missing values
# separate EV/WCL exclusive proteins (don't impute) from 
# EV/WCL common proteins (impute)
yeast_excl <- filter_na2(yeast_norm2, logic = "or", op = "==", 
                         pat1 = "EV", val1 = 3, 
                         pat2 = "W", val2 = 3)

@@

# proteins with missing values tend to have lower intensity
# therefore proteins are MNAR, close to detection limit
# use left censored imputation method
yeast_imp <- impute_QRILC(yeast_both)

@@

# extract DE results
yeast_res <- get_results(efit = y_efit, 
                         mat = yeast_de, 
                         p_val = 0.01, 
                         lfc = 0, 
                         type = "individual")[[1]]
```

@@

# can't impute with so many missing values
# filter for proteins with 4-5 NA values in EV or WCL (don't impute)
biofilm_excl <- filter_na2(biofilm_norm, logic = "or", op = ">=", 
                         pat1 = "EV", val1 = 4, 
                         pat2 = "W", val2 = 4)

@@

# proteins with missing values tend to have lower intensity
# therefore proteins are MNAR, close to detection limit
# use left censored imputation method
biofilm_imp <- impute_QRILC(biofilm_both)

@@

# extract DE results
biofilm_res <- get_results(efit = b_efit, 
                           mat = biofilm_de, 
                           p_val = 0.01, 
                           lfc = 0, 
                           type = "individual")[[1]]
```

@@

```{r atcc1, include=FALSE}
# filter for proteins quantified in min 2/3 reps of EV or WCL
a1_norm <- filter_na2(atcc_norm, logic = "or", op = "<=", 
                      pat1 = "A10231_EV", val1 = 1, 
                      pat2 = "A10231_W", val2 = 1)
a1_norm <- a1_norm[, -c(7:12)]

@@

# can't impute with so many missing values
# separate EV/WCL exclusive proteins (don't impute) from
# EV/WCL common proteins (impute)
a1_excl <- filter_na2(a1_norm, logic = "or", op = "==",
                      pat1 = "EV", val1 = 3,
                      pat2 = "W", val2 = 3)

@@

# proteins with missing values tend to have lower intensity
# therefore proteins are MNAR, close to detection limit
# use left censored imputation method
a1_imp <- impute_QRILC(a1_both)

@@

# extract DE results
a1_res <- get_results(efit = a1_efit,
                      mat = a1_de,
                      p_val = 0.01,
                      lfc = 0,
                      type = "individual")[[1]]
```

```{r atcc9, include=FALSE}
# filter for proteins quantified in min 2/3 reps of EV or WCL
a9_norm <- filter_na2(atcc_norm, logic = "or", op = "<=",
                      pat1 = "A90028_EV", val1 = 1,
                      pat2 = "A90028_W", val2 = 1)
a9_norm <- a9_norm[, -c(1:6)]

@@

# proteins with missing values tend to have lower intensity
# therefore proteins are MNAR, close to detection limit
# use left censored imputation method
a9_imp <- impute_QRILC(a9_both)

@@

# extract DE results
a9_res <- get_results(efit = a9_efit,
                      mat = a9_de,
                      p_val = 0.01,
                      lfc = 0,
                      type = "individual")[[1]]
```

@@

# construct table 1
marker_longlist <- marker_candidates %>% 
  mutate(Function = match_id(Name, uniprot, "CGD_gene_name", "Protein_name"),
         DAY.Y = match_id(Name, yeast_res, "CGD_gene_name", "logFC", 
                          str_split = FALSE), 
         A9 = match_id(Name, a9_res, "CGD_gene_name", "logFC", 
                       str_split = FALSE), 
         A1 = match_id(Name, a1_res, "CGD_gene_name", "logFC", 
                       str_split = FALSE), 
         DAY.B = match_id(Name, biofilm_res, "CGD_gene_name", "logFC", 
                          str_split = FALSE), 
         TM = match_id(Name, uniprot, "CGD_gene_name", "Transmembrane"), 
         SP = match_id(Name, uniprot, "CGD_gene_name", "Signal_peptide")) %>% 
  mutate_at(.vars = vars(DAY.Y, DAY.B, A1, A9), 
            .funs = list(~ sprintf(as.numeric(.), fmt = "%#.2f"))) %>% 
  mutate_at(.vars = vars(DAY.Y, DAY.B, A1, A9), 
            .funs = list(~ case_when(. == "NA" ~ "ex", 
                                     TRUE ~ as.character(.)))) %>% 
  mutate(TM = stringr::str_count(TM, "Helical"), 
         SP = case_when(grepl("SIGNAL", SP) ~ "Y", 
                        TRUE ~ ""), 
         TM = case_when(TM == 0 ~ "", 
                        TRUE ~ as.character(TM))) %>% 
  rename("DAY Y" = DAY.Y, "DAY B" = DAY.B) %>% 
  arrange(Name)

# load vector of protein functions from external R script
<<fun_list1>>

# add proper protein functions to table 1
# also add vesicle depleted supernatant info and 
# group proteins according to localisation
marker_longlist2 <- marker_longlist %>% 
  mutate(Function = function_desc1, 
         pack_row = c(4, rep(1, 3), rep(4, 2), 1, 2, rep(1, 6), 3, rep(4, 2),
                      1, rep(2, 3), 1, rep(4, 2), 1, rep(4, 3), 1, 3, 
                      rep(1, 3), 4, 1, 2, 1, 2, 4)) %>% 
  arrange(pack_row) %>% 
  mutate(VDM = c(rep("", 11), "Y", rep("", 7), "Y", rep("", 7), 
                 rep("Y", 7), "", rep("Y", 4)))

# fix number of TM domains for CDR1;CDR2
marker_longlist2[2, 7] <- "12;12"

# UniProt says SUR7 has 3 TM domains and a signal peptide
# much of literature says SUR7 has 4 TM domains
# TOPCONS2 predicts SUR7 has 4 TM domains and no signal peptide
# fix SUR7 entries
marker_longlist2[18, 7] <- "4"
marker_longlist2[18, 8] <- ""

# Change name of orf19.6741 to Evp1
marker_longlist2[12, 1] <- paste0("EVP1",
                                  footnote_marker_alphabet(4))
```

<!-- assign markdown references for table caption -->
(ref:table1) [@gil-bona2015]
(ref:footnote1) [@skrzypek2017]
(ref:footnote2) [@almagroarmenteros2017]
(ref:footnote3) [@tsirigos2015]

```{r table1}
tab1_cap <- "\\textbf{Candidate positive protein markers for 
             \\textit{C. albicans} EVs.} This list of proteins consists of 
             those that were found to be exclusive to EVs or significantly 
             enriched in EVs across the four \\textit{C. albicans} strains 
             examined in this study. Proteins are grouped according to their 
             subcellular localisation as annotated in the Candida Genome 
             Database (candidagenome.org) (ref:footnote1). 
             The log\\textsubscript{2} ratio of the abundance (mean MaxQuant 
             LFQ intensity) of each protein in EVs compared to whole cell 
             lysate (WCL) for each strain is listed. \`\`ex\" indicates where 
             a protein was only quantified in the EV fraction and not the WCL 
             for that strain. The \`\`TM\" column indicates the number of 
             transmembrane domains for each protein as annotated in UniProtKB. 
             \`\`SP\" indicates whether a protein is annotated as having a 
             signal peptide according to UniProtKB. \`\`VDM\" shows whether a 
             protein has previously been detected in vesicle-depleted culture 
             media (i.e. the proteins may also be in the soluble secretome) 
             (ref:table1). Underlined proteins are those identified as the 
             best candidates for positive EV markers according to the 
             criteria depicted in supplemental Figure S1."

# draw table 1
knitr::kable(marker_longlist2[ ,-9],
             caption = tab1_cap, 
             format = "latex", 
             booktabs = TRUE, 
             longtable = TRUE, 
             linesep = "", 
             escape = FALSE)  %>% 
  row_spec(row = c(1, 5, 11, 12, 15:19, 21:25, 35), underline = TRUE) %>% 
  add_header_above(c(" " = 2, 
                     "log\\\\textsubscript{2}(fold change) EV vs WCL" = 4, 
                     " " = 2), escape = FALSE) %>% 
  kable_styling(font_size = 8) %>% 
  pack_rows("Plasma membrane", 1, 19) %>% 
  pack_rows("Golgi, ER, vacuole", 20, 25) %>% 
  pack_rows("Mitochondria", 26, 27) %>% 
  pack_rows("Cell wall, cell surface", 28, 39) %>% 
  footnote(alphabet = c("Protein localisation was inferred from sequence 
                         similarity with \\\\textit{S. cerevisiae} ortholog as 
                         annotated in the Candida Genome Database 
                        (ref:footnote1).", 
                        "Protein localisation was predicted using DeepLoc-1.0 
                        (ref:footnote2).", 
                        "Presence of transmembrane domains and absense of a 
                        signal peptide was predicted using TOPCONS2 
                        (ref:footnote3).",
                        "The name Evp1 for the protein encoded by orf19.6741 
                        was proposed in the present study."), 
           escape = FALSE, 
           threeparttable = TRUE)
```

@@

# generate volcano plot
f2b1 <- plot_volcano(data = yeast_res, 
                     p_val = adj.P.Val, 
                     log2fc = logFC,
                     group = group, 
                     use_labels = TRUE,
                     label_fun = ifelse(logFC >= 8, CGD_gene_name, ""), 
                     point_size = 0.6, 
                     label_size = 2, 
                     x_lim = c(-10, 12), 
                     y_lim = c(0, 8), 
                     legend_params = c("left", 11, 8), 
                     axis_params = c("black", 10)) + 
  scale_colour_manual(labels = c("EV enriched", 
                                 "WCL enriched", 
                                 "Not significant"), 
                      values = unname(ev_pal[c(2,1,3)]), 
                      breaks = c("EV up", "WCL up", "not sig"), 
                      name = element_blank()) + 
  annotate("text", x = 8.5, y = 0.5, label = "bold('DAY286 Y')", parse = TRUE) + 
  annotate("text", x = 3.5, y = 7, label = "bold('88')", 
           parse = TRUE, colour = unname(ev_pal[2])) + 
  annotate("text", x = -3.5, y = 7, label = "bold('88')", parse = TRUE, 
           colour = unname(ev_pal[3])) + 
  annotate("text", x = 0, y = 6, label = "bold('318')", parse = TRUE, 
           colour = unname(ev_pal[1]))
```

@@

# generate volcano plot
f2b2 <- plot_volcano(data = a9_res, 
                     p_val = adj.P.Val, 
                     log2fc = logFC,
                     group = group, 
                     use_labels = TRUE, 
                     label_fun = ifelse(logFC >= 8, CGD_gene_name, ""), 
                     point_size = 0.6, 
                     label_size = 2, 
                     x_lim = c(-10, 12), 
                     y_lim = c(0, 8), 
                     legend_params = c("left", 11, 8), 
                     axis_params = c("black", 10)) + 
  scale_colour_manual(labels = c("EV enriched (86)", "WCL enriched (122)", 
                                 "Not significant (761)"),
                      values = unname(ev_pal[c(2,3,1)])) + 
  annotate("text", x = 8, y = 0.5, label = "bold('ATCC90028')", parse = TRUE) + 
  annotate("text", x = 3.5, y = 7, label = "bold('86')", 
           parse = TRUE, colour = unname(ev_pal[2])) + 
  annotate("text", x = -3.5, y = 7, label = "bold('122')", 
           parse = TRUE, colour = unname(ev_pal[3])) + 
  annotate("text", x = 0, y = 6, label = "bold('761')", 
           parse = TRUE, colour = unname(ev_pal[1]))
```

@@

# generate volcano plot
f2b3 <- plot_volcano(data = a1_res, 
                     p_val = adj.P.Val, 
                     log2fc = logFC,
                     group = group, 
                     use_labels = TRUE, 
                     label_fun = ifelse(logFC >= 8, CGD_gene_name, ""), 
                     point_size = 0.6, 
                     label_size = 2, 
                     x_lim = c(-10, 12), 
                     y_lim = c(0, 8), 
                     legend_params = c("left", 11, 8), 
                     axis_params = c("black", 10)) + 
  scale_colour_manual(labels = c("EV enriched (55)", "WCL enriched (52)", 
                                 "Not significant (567)"),
                      values = unname(ev_pal[c(2,3,1)])) + 
  annotate("text", x = 8, y = 0.5, label = "bold('ATCC10231')", parse = TRUE) + 
  annotate("text", x = 3.5, y = 7, label = "bold('55')", 
           parse = TRUE, colour = unname(ev_pal[2])) + 
  annotate("text", x = -3.5, y = 7, label = "bold('52')", 
           parse = TRUE, colour = unname(ev_pal[3])) + 
  annotate("text", x = 0, y = 6, label = "bold('567')", 
           parse = TRUE, colour = unname(ev_pal[1]))
```

@@

# generate volcano plot
f2b4 <- plot_volcano(data = biofilm_res, 
                     p_val = adj.P.Val, 
                     log2fc = logFC,
                     group = group, 
                     use_labels = TRUE, 
                     label_fun = ifelse(logFC >= 8, CGD_gene_name, ""), 
                     point_size = 0.6, 
                     label_size = 2, 
                     x_lim = c(-10, 12), 
                     y_lim = c(0, 8), 
                     legend_params = c("left", 11, 8), 
                     axis_params = c("black", 10)) + 
  scale_colour_manual(labels = c("EV enriched (95)", "WCL enriched (79)", 
                                 "Not significant (484)"), 
                      values = unname(ev_pal[c(2,1,3)]), 
                      breaks = c("EV up", "WCL up", "not sig")) + 
  annotate("text", x = 8.5, y = 0.5, label = "bold('DAY286 B')", parse = TRUE) + 
  annotate("text", x = 3.5, y = 7, label = "bold('95')", 
           parse = TRUE, colour = unname(ev_pal[2])) + 
  annotate("text", x = -3.5, y = 7, label = "bold('79')", 
           parse = TRUE, colour = unname(ev_pal[3])) + 
  annotate("text", x = 0, y = 6, label = "bold('484')", 
           parse = TRUE, colour = unname(ev_pal[1]))
```

@@

f2_cap <- "\\textbf{Differential abundance analysis of proteins identified in 
           \\textit{C. albicans} EVs and whole cell lysates (WCL).} 
           \\textit{(A)} Venn diagrams comparing the EV and WCL proteomes from 
           four \\textit{C. albicans} strains; DAY286 yeast (n = 3), ATCC90028 
           yeast (n = 3), ATCC10231 yeast (n = 3), and DAY286 biofilm (n = 5). 
           Whole cell lysates were prepared from the EV source cells. Proteins 
           were quantified (MaxQuant LFQ intensity) in minimum 2/3 (yeast) and 
           4/5 (biofilm) biological replicates of EV or WCL. \\textit{(B)} 
           Volcano plots depicting significantly enriched EV or WCL proteins. 
           Differential abundance analysis was performed by comparing the mean 
           normalised LFQ intensities of proteins identified in both EV and 
           WCL (i.e. proteins in the venn overlap) using the Bioconductor 
           package \\textit{limma} [@cox2014; @ritchie2015]. Significantly 
           enriched proteins were identified using a Benjamini-Hochberg 
           adjusted p-value cut off of 0.01. Counts of significant and 
           non-significant proteins are indicated on each graph. Proteins with 
           a log\\textsubscript{2}(FC) greater than 8 are labelled. Data tables 
           underlying the venn diagrams and volcano plots are provided in 
           supplemental File S2."
```

@@

# figure 4c MDS [dim1, dim2]
f4c <- plot_mds(mds_data, mat_labels = mds_labels, 
                shape_size = 3, dim.plot = c(1,2)) + 
  scale_x_continuous(limits = c(-1.5, 2.5), 
                     expand = c(0,0)) + 
  scale_y_continuous(limits = c(-2.5, 1.5), 
                     expand = c(0,0)) + 
  scale_fill_manual(breaks = c("Y", "A9", "A1", "B"), 
                    values = st_pal[c(3, 2, 4, 1)], 
                    labels = c("DAY Y", "A90028", "A10231", "DAY B"), 
                    guide = "legend") + 
  scale_shape_manual(values = c(22,24)) + 
  theme_bw() + 
  theme(legend.title = element_blank(),
        legend.position = "right", 
        legend.text = element_text(size = 8), 
        legend.direction = "vertical") + 
  guides(fill = guide_legend(override.aes = list(shape = 21))) + 
  xlab(label = "Dimension 1") + 
  ylab(label = "Dimension 2")

# figure 4d MDS [dim3, dim2]
f4d <- plot_mds(mds_data, mat_labels = mds_labels, 
                shape_size = 3, dim.plot = c(3,2)) + 
  scale_x_continuous(limits = c(-1.5, 2.5), 
                     expand = c(0,0)) + 
  scale_y_continuous(limits = c(-2.5, 1.5), 
                     expand = c(0,0)) + 
  scale_fill_manual(breaks = c("Y", "A9", "A1", "B"), 
                    values = st_pal[c(3, 2, 4, 1)], 
                    labels = c("DAY Y", "A90028", "A10231", "DAY B"), 
                    guide = "legend") + 
  scale_shape_manual(values = c(22,24)) + 
  theme_bw() + 
  theme(legend.title = element_blank(),
        legend.position = "right", 
        legend.text = element_text(size = 8), 
        legend.direction = "vertical", 
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0.1), 
        legend.box.margin = margin(t = -10, r = -6, b = -10, l = -10)) + 
  guides(fill = guide_legend(override.aes = list(shape = 21))) + 
  xlab(label = "Dimension 3") + 
  ylab(label = "Dimension 2")
```

>>>>

@@

# won't impute with so many missing values
# separate EV/WCL exclusive proteins (don't impute) from 
# EV/WCL common proteins (impute)
yeast_excl <- filter_na2(yeast_norm2, logic = "or", op = "==", 
                         pat1 = "EV", val1 = 3, 
                         pat2 = "W", val2 = 3)

@@

# use bpca method from pcaMethods for missing value imputation
yeast_imp <- pcaMethods::pca(yeast_both, method = "bpca")
yeast_imp <- pcaMethods::completeObs(yeast_imp)

@@

# extract DE results
# adj.p.val cutoff is 0.01, log2FC cutoff is 1
yeast_res <- get_results(efit = y_efit, 
                         mat = yeast_de, 
                         p_val = 0.01, 
                         lfc = 1, 
                         type = "individual")[[1]]
```

@@

# won't impute with so many missing values
# filter for proteins with 4-5 NA values in EV or WCL (don't impute)
biofilm_excl <- filter_na2(biofilm_norm, logic = "or", op = ">=", 
                         pat1 = "EV", val1 = 4, 
                         pat2 = "W", val2 = 4)

@@

# use bpca method from pcaMethods for missing value imputation
biofilm_imp <- pcaMethods::pca(biofilm_both, method = "bpca")
biofilm_imp <- pcaMethods::completeObs(biofilm_imp)

@@

# extract DE results
# adj.p.val cutoff is 0.01, log2FC cutoff is 1
biofilm_res <- get_results(efit = b_efit, 
                           mat = biofilm_de, 
                           p_val = 0.01, 
                           lfc = 1, 
                           type = "individual")[[1]]
```

@@

```{r atcc1, include=FALSE}
# filter for proteins quantified in min 2/3 reps of ATCC10231 EV or WCL
a1_norm <- filter_na2(atcc_norm, logic = "or", op = "<=", 
                      pat1 = "A10231_EV", val1 = 1, 
                      pat2 = "A10231_W", val2 = 1)
a1_norm <- a1_norm[, -c(7:12)]

@@

# won't impute with so many missing values
# separate EV/WCL exclusive proteins (don't impute) from
# EV/WCL common proteins (impute)
a1_excl <- filter_na2(a1_norm, logic = "or", op = "==",
                      pat1 = "EV", val1 = 3,
                      pat2 = "W", val2 = 3)

@@

# use bpca method from pcaMethods for missing value imputation
a1_imp <- pcaMethods::pca(a1_both, method = "bpca")
a1_imp <- pcaMethods::completeObs(a1_imp)

@@

# extract DE results
# adj.p.val cutoff is 0.01, log2FC cutoff is 1
a1_res <- get_results(efit = a1_efit,
                      mat = a1_de,
                      p_val = 0.01,
                      lfc = 1,
                      type = "individual")[[1]]
```

```{r atcc9, include=FALSE}
# filter for proteins quantified in min 2/3 reps of ATCC90028 EV or WCL
a9_norm <- filter_na2(atcc_norm, logic = "or", op = "<=",
                      pat1 = "A90028_EV", val1 = 1,
                      pat2 = "A90028_W", val2 = 1)
a9_norm <- a9_norm[, -c(1:6)]

@@

# use bpca method from pcaMethods for missing value imputation
a9_imp <- pcaMethods::pca(a9_both, method = "bpca")
a9_imp <- pcaMethods::completeObs(a9_imp)

@@

# extract DE results
# adj.p.val cutoff is 0.01, log2FC cutoff is 1
a9_res <- get_results(efit = a9_efit,
                      mat = a9_de,
                      p_val = 0.01,
                      lfc = 1,
                      type = "individual")[[1]]
```

@@

# construct table 1
marker_longlist <- marker_candidates %>% 
  mutate(DAY.Y = match_id(Name, yeast_res, "CGD_gene_name", "logFC", 
                          str_split = FALSE), 
         A9 = match_id(Name, a9_res, "CGD_gene_name", "logFC", 
                       str_split = FALSE), 
         A1 = match_id(Name, a1_res, "CGD_gene_name", "logFC", 
                       str_split = FALSE), 
         DAY.B = match_id(Name, biofilm_res, "CGD_gene_name", "logFC", 
                          str_split = FALSE), 
         TM = match_id(Name, uniprot, "CGD_gene_name", "Transmembrane"), 
         SP = match_id(Name, uniprot, "CGD_gene_name", "Signal_peptide")) %>% 
  mutate_at(.vars = vars(DAY.Y, DAY.B, A1, A9), 
            .funs = list(~ sprintf(as.numeric(.), fmt = "%#.2f"))) %>% 
  mutate_at(.vars = vars(DAY.Y, DAY.B, A1, A9), 
            .funs = list(~ case_when(. == "NA" ~ "ex", 
                                     TRUE ~ as.character(.)))) %>% 
  mutate(TM = stringr::str_count(TM, "Helical"), 
         SP = case_when(grepl("SIGNAL", SP) ~ "Y", 
                        TRUE ~ ""), 
         TM = case_when(TM == 0 ~ "", 
                        TRUE ~ as.character(TM))) %>% 
  rename("DAY Y" = DAY.Y, "DAY B" = DAY.B) %>% 
  arrange(Name)

# load .csv file with table 1 labels
marker_labels <- read.csv("components/table1.csv", stringsAsFactors = FALSE) %>% 
  mutate(Function = case_when(Superscript == "a" ~ paste0(Function, footnote_marker_alphabet(1)), 
                              Superscript == "b"~ paste0(Function, footnote_marker_alphabet(2)), 
                              Superscript == "c" ~ paste0(Function, footnote_marker_alphabet(3)), 
                              TRUE ~ as.character(Function)))

# add proper protein functions to table 1
# also add vesicle depleted supernatant info and 
# group proteins according to localisation
marker_longlist2 <- match_id_multi(df = marker_longlist, 
                                   id = "Name", 
                                   ref = marker_labels, 
                                   match = "Name", 
                                   new = c("Function", "Loc_num", "VDM"), 
                                   str_split = FALSE) %>% 
  select(Name, Function, everything()) %>% 
  arrange(Loc_num)

# fix number of TM domains for CDR1;CDR2
marker_longlist2[3, 7] <- "12;12"

# UniProt says SUR7 has 3 TM domains and a signal peptide
# much of literature says SUR7 has 4 TM domains
# TOPCONS2 predicts SUR7 has 4 TM domains and no signal peptide
# fix SUR7 entries
marker_longlist2[22, 7] <- "4"
marker_longlist2[22, 8] <- ""

# Change name of orf19.6741 to Evp1
marker_longlist2[15, 1] <- paste0("EVP1",
                                  footnote_marker_alphabet(4))
```

<!-- assign markdown references for table caption -->
(ref:table1) [@gil-bona2015]
(ref:footnote1) [@skrzypek2017]
(ref:footnote2) [@theuniprotconsortium2019]
(ref:footnote3) [@tsirigos2015]

```{r table1}
tab1_cap <- "\\textbf{Candidate positive protein markers for 
             \\textit{C. albicans} EVs.} This list of proteins consists of 
             those that were found to be exclusive to EVs or significantly 
             enriched in EVs across the four \\textit{C. albicans} strains 
             examined in this study. Proteins are grouped according to their 
             subcellular localisation as annotated in the Candida Genome 
             Database (candidagenome.org) (ref:footnote1) or in the 
             \\textit{C. albicans} UniProt reference proteome (UP000000559) 
             (ref:footnote2). The log\\textsubscript{2} ratio of the abundance 
             (mean MaxQuant LFQ intensity) of each protein in EVs compared to 
             whole cell lysate (WCL) for each strain is listed. \`\`ex\" 
             indicates where a protein was only quantified in the EV fraction 
             and not the WCL for that strain. The \`\`TM\" column indicates the 
             number of transmembrane domains for each protein as annotated in 
             UniProt. \`\`SP\" indicates whether a protein is annotated as 
             having a signal peptide according to UniProt. \`\`VDM\" shows 
             whether a protein has previously been detected in vesicle-depleted 
             culture media (i.e. the proteins may also be in the soluble secretome) 
             (ref:table1). Underlined proteins are those identified as the 
             best candidates for positive EV markers according to the 
             criteria depicted in supplemental Figure S1."

# draw table 1
knitr::kable(marker_longlist2[ ,-9],
             caption = tab1_cap, 
             format = "latex", 
             booktabs = TRUE, 
             longtable = TRUE, 
             linesep = "", 
             escape = FALSE)  %>% 
  row_spec(row = c(1,2, 6,7, 13,14,15, 18,19,20,21,22,23, 27, 32, 
                   38,39,40, 42,43,44,45), underline = TRUE) %>% 
  add_header_above(c(" " = 2, 
                     "log\\\\textsubscript{2}(fold change) EV vs WCL" = 4, 
                     " " = 2), escape = FALSE) %>% 
  kable_styling(font_size = 8) %>% 
  pack_rows("Plasma membrane", 1, 23) %>% 
  pack_rows("Cell wall, cell surface", 24, 36) %>% 
  pack_rows("Endoplasmic reticulum", 37, 41) %>% 
  pack_rows("Endosome, Golgi, transport vesicle", 42, 44) %>% 
  pack_rows("Vacuole", 45, 45) %>% 
  pack_rows("Mitochondrion", 46, 47) %>% 
  footnote(alphabet = c("Protein localisation was inferred from sequence 
                         similarity with \\\\textit{S. cerevisiae} ortholog as 
                         annotated in the Candida Genome Database 
                        (ref:footnote1).", 
                        "Protein localisation was obtained from the GO Cellular
                        Component annotation in UniProt (ref:footnote2).", 
                        "Presence of transmembrane domains and absence of a 
                        signal peptide was predicted using TOPCONS2 
                        (ref:footnote3).",
                        "The name Evp1 for the protein encoded by orf19.6741 
                        was proposed in the present study."), 
           escape = FALSE, 
           threeparttable = TRUE)
```

@@

# generate volcano plot
f2b1 <- plot_volcano(data = yeast_res, 
                     p_val = adj.P.Val, 
                     log2fc = logFC,
                     group = group, 
                     use_labels = TRUE,
                     label_fun = ifelse(logFC >= 8, CGD_gene_name, ""), 
                     point_size = 0.6, 
                     label_size = 2, 
                     x_lim = c(-10, 12), 
                     y_lim = c(0, 8), 
                     legend_params = c("left", 11, 8), 
                     axis_params = c("black", 10)) + 
  scale_colour_manual(labels = c("EV enriched", 
                                 "WCL enriched", 
                                 "Not significant"), 
                      values = unname(ev_pal[c(2,1,3)]), 
                      breaks = c("EV up", "WCL up", "not sig"), 
                      name = element_blank()) + 
  annotate("text", x = 8.5, y = 0.5, label = "bold('DAY286 Y')", parse = TRUE) + 
  annotate("text", x = 3.5, y = 7, label = "bold('101')", 
           parse = TRUE, colour = unname(ev_pal[2])) + 
  annotate("text", x = -3.5, y = 7, label = "bold('117')", parse = TRUE, 
           colour = unname(ev_pal[3])) + 
  annotate("text", x = 0, y = 6, label = "bold('276')", parse = TRUE, 
           colour = unname(ev_pal[1]))
```

@@

# generate volcano plot
f2b2 <- plot_volcano(data = a9_res, 
                     p_val = adj.P.Val, 
                     log2fc = logFC,
                     group = group, 
                     use_labels = TRUE, 
                     label_fun = ifelse(logFC >= 8, CGD_gene_name, ""), 
                     point_size = 0.6, 
                     label_size = 2, 
                     x_lim = c(-10, 12), 
                     y_lim = c(0, 8), 
                     legend_params = c("left", 11, 8), 
                     axis_params = c("black", 10)) + 
  scale_colour_manual(labels = c("EV enriched", "WCL enriched", 
                                 "Not significant"),
                      values = unname(ev_pal[c(2,3,1)])) + 
  annotate("text", x = 8, y = 0.5, label = "bold('ATCC90028')", parse = TRUE) + 
  annotate("text", x = 3.5, y = 7, label = "bold('121')", 
           parse = TRUE, colour = unname(ev_pal[2])) + 
  annotate("text", x = -3.5, y = 7, label = "bold('176')", 
           parse = TRUE, colour = unname(ev_pal[3])) + 
  annotate("text", x = 0, y = 6, label = "bold('672')", 
           parse = TRUE, colour = unname(ev_pal[1]))
```

@@

# generate volcano plot
f2b3 <- plot_volcano(data = a1_res, 
                     p_val = adj.P.Val, 
                     log2fc = logFC,
                     group = group, 
                     use_labels = TRUE, 
                     label_fun = ifelse(logFC >= 8, CGD_gene_name, ""), 
                     point_size = 0.6, 
                     label_size = 2, 
                     x_lim = c(-10, 12), 
                     y_lim = c(0, 8), 
                     legend_params = c("left", 11, 8), 
                     axis_params = c("black", 10)) + 
  scale_colour_manual(labels = c("EV enriched", "WCL enriched", 
                                 "Not significant"),
                      values = unname(ev_pal[c(2,3,1)])) + 
  annotate("text", x = 8, y = 0.5, label = "bold('ATCC10231')", parse = TRUE) + 
  annotate("text", x = 3.5, y = 7, label = "bold('101')", 
           parse = TRUE, colour = unname(ev_pal[2])) + 
  annotate("text", x = -3.5, y = 7, label = "bold('105')", 
           parse = TRUE, colour = unname(ev_pal[3])) + 
  annotate("text", x = 0, y = 6, label = "bold('468')", 
           parse = TRUE, colour = unname(ev_pal[1]))
```

@@

# generate volcano plot
f2b4 <- plot_volcano(data = biofilm_res, 
                     p_val = adj.P.Val, 
                     log2fc = logFC,
                     group = group, 
                     use_labels = TRUE, 
                     label_fun = ifelse(logFC >= 8, CGD_gene_name, ""), 
                     point_size = 0.6, 
                     label_size = 2, 
                     x_lim = c(-10, 12), 
                     y_lim = c(0, 8), 
                     legend_params = c("left", 11, 8), 
                     axis_params = c("black", 10)) + 
  scale_colour_manual(labels = c("EV enriched", "WCL enriched", 
                                 "Not significant"), 
                      values = unname(ev_pal[c(2,1,3)]), 
                      breaks = c("EV up", "WCL up", "not sig")) + 
  annotate("text", x = 8.5, y = 0.5, label = "bold('DAY286 B')", parse = TRUE) + 
  annotate("text", x = 3.5, y = 7, label = "bold('105')", 
           parse = TRUE, colour = unname(ev_pal[2])) + 
  annotate("text", x = -3.5, y = 7, label = "bold('100')", 
           parse = TRUE, colour = unname(ev_pal[3])) + 
  annotate("text", x = 0, y = 6, label = "bold('453')", 
           parse = TRUE, colour = unname(ev_pal[1]))
```

@@

f2_cap <- "\\textbf{Differential abundance analysis of proteins identified in 
           \\textit{C. albicans} EVs and whole cell lysates (WCL).} 
           \\textit{(A)} Venn diagrams comparing the EV and WCL proteomes from 
           four \\textit{C. albicans} strains; DAY286 yeast (n = 3), ATCC90028 
           yeast (n = 3), ATCC10231 yeast (n = 3), and DAY286 biofilm (n = 5). 
           Whole cell lysates were prepared from the EV source cells. Proteins 
           were quantified (MaxQuant LFQ intensity) in minimum 2/3 (yeast) and 
           4/5 (biofilm) biological replicates of EV or WCL. \\textit{(B)} 
           Volcano plots depicting significantly enriched EV or WCL proteins. 
           Differential abundance analysis was performed by comparing the mean 
           normalised LFQ intensities of proteins identified in both EV and 
           WCL (i.e. proteins in the venn overlap) using the Bioconductor 
           package \\textit{limma} [@cox2014; @ritchie2015]. Significantly 
           enriched proteins were identified using a Benjamini-Hochberg 
           adjusted p-value cut off of 0.01 and log\\textsubscript{2}(FC) cutoff 
           of 1. Counts of significant and 
           non-significant proteins are indicated on each graph. Proteins with 
           a log\\textsubscript{2}(FC) greater than 8 are labelled. Data tables 
           underlying the venn diagrams and volcano plots are provided in 
           supplemental File S2."
```

@@

# figure 4c MDS [dim1, dim2]
f4c <- plot_mds(mds_data, mat_labels = mds_labels, 
                shape_size = 3, dim.plot = c(1,2)) + 
  scale_x_continuous(limits = c(-1.5, 2), 
                     expand = c(0,0)) + 
  scale_y_continuous(limits = c(-2, 1.5), 
                     expand = c(0,0)) + 
  scale_fill_manual(breaks = c("Y", "A9", "A1", "B"), 
                    values = st_pal[c(3, 2, 4, 1)], 
                    labels = c("DAY Y", "A90028", "A10231", "DAY B"), 
                    guide = "legend") + 
  scale_shape_manual(values = c(22,24)) + 
  theme_bw() + 
  theme(legend.title = element_blank(),
        legend.position = "right", 
        legend.text = element_text(size = 8), 
        legend.direction = "vertical") + 
  guides(fill = guide_legend(override.aes = list(shape = 21))) + 
  xlab(label = "Dimension 1") + 
  ylab(label = "Dimension 2")

# figure 4d MDS [dim3, dim2]
f4d <- plot_mds(mds_data, mat_labels = mds_labels, 
                shape_size = 3, dim.plot = c(3,2)) + 
  scale_x_continuous(limits = c(-1.5, 2), 
                     expand = c(0,0)) + 
  scale_y_continuous(limits = c(-2, 1.5), 
                     expand = c(0,0)) + 
  scale_fill_manual(breaks = c("Y", "A9", "A1", "B"), 
                    values = st_pal[c(3, 2, 4, 1)], 
                    labels = c("DAY Y", "A90028", "A10231", "DAY B"), 
                    guide = "legend") + 
  scale_shape_manual(values = c(22,24)) + 
  theme_bw() + 
  theme(legend.title = element_blank(),
        legend.position = "right", 
        legend.text = element_text(size = 8), 
        legend.direction = "vertical", 
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0.1), 
        legend.box.margin = margin(t = -10, r = -6, b = -10, l = -10)) + 
  guides(fill = guide_legend(override.aes = list(shape = 21))) + 
  xlab(label = "Dimension 3") + 
  ylab(label = "Dimension 2")
```
